---
title: "R graphics dummies"
author: "Kenny"
date: "Thursday, January 22, 2015"
output: html_document
---


### Graphics

```{r}
x <- 1:5                     # 5 x-axis points
y <- x^2                     # 5 y-axis points
plot(x, y, pch=3:7)          # 5 points = 5 symbols
plot(x, y, pch=c("a", "b"))  # 5 points = 2 symbols cycling to match 
plot(x, y, col="red")        # Each symbol is red
plot(x, y, pch=c("a", "b"),  # Symbols are a b cycling to match 5 points
     col=c("red", "blue", "green")) # 3 cols cycling for plotting color = border color
# Plot circles with a thick red border and blue fill color
plot(x, y, pch=21, lwd=3, col="red", bg="blue") # bg = background color = fill color
plot(x, y, cex=seq(1,2,.25)) # cex = size, cex = a vector of 5 different sizes, increasing
plot(x, y, type="l", lty=2)  # type = 'l' to change circle to line,  line type = pattern 2
plot(x, y, type="l", lwd=3)  # default line width = 1, here line width 3
par(lwd=3, pch=4, cex.main = 3)  # part()go before plot(), set up Line width, symbol, cex
plot(x, y, lwd=1, main = "my chart") # write main = title string here
par(cex.lab=2, cex.main=3)   # in par() set label size with cex.lab
plot(x, y, xlim=c(0, 10))    # xlim = set x axis from 0 to 10
plot(x, y, log="y")          # do Logarithmic to y axis data
plot(x, y, las=1)            # las can be 0-3, but no idea what difference to make

text(1, 2, "Where's Wally?")  # Add text to existing plot, check ?text
mtext("I'm out here", side=4) # and text to the right margin, ?mtext
```

### math expressions in R text
```{r}
demo(plotmath)
```


### Including Greek letters and equations in graphs

```{r}
x <- seq(0, 5, .1)
plot(x, sin(x), type="l", ylim=c(-1, 2.5),
     xlab=expression(paste("Concentration ", mu[i])))
lines(x, sqrt(x), lty=2)
title(expression(paste("This looks like ", Gamma, " to me")))

text(4, 1.5, expression(hat(x) == sqrt(x)))
     
```

> only the following code has something wrong  

```
text(1.6, -.5, expression(paste(plain(sin)(x) == 
                                sum(frac((-1)^n, paste((2*n+1), plain("!")))*x^(2*n+1), 
                                      n==0, infinity)))
```


### Set colors in R graphics

```{r}
colors()  # show all the colors in R system
plot(1:10, col="cyan")         # 1:10 apply to both x and y; Plot using cyan color
plot(1:10, col="#cc0076")      # Plot using purplish color

rgb(10, 20, 30, maxColorValue=255) # returns a code for color like "#0A141E"
plot(1:10, col=rgb(10, 20, 30, maxColorValue=255))      # Plot using purplish color

plot(1:10, col="#0A141E")      # Plot using purplish color

rgb(10, 20, 30, alpha=50, maxColorValue=255) # alpha set the level of transparency compare to the max number

rgb(10, 20, 30, maxColorValue=100)
rgb(10, 20, 30, alpha=50, maxColorValue=100)
```


### Set color palettes in R graphics

```{r}
palette()             # Show the current palette in R 
palette(c("red", "#1A334D", "black", rgb(.1, .8, 0))) # set the palette with specific color you want
palette() # display the specific color you set
palette("default")    # Use the default palette
library(RColorBrewer) # Full color palettes can be generated directly without having to specify each color separately
brewer.pal.info
str(brewer.pal.info)
head(brewer.pal.info, n=10) # show only ten of those color palette names above
brewer.pal(n=5, name="Spectral")       # choose 5 colors from Spectral palette
brewer.pal(n=10, name="Accent")        # 10 colors from Accent
# Set colors locally for just one function call
plot(1:8, col=brewer.pal(n=8, name="Reds"), pch=16, cex=4) # give each point a color of the eight
palette(brewer.pal(n=8, name="Greens")) # Set global palette for the plotting following
plot(1:8, col=1:8, pch=16, cex=4)       # Plot using colors

```


### High-level plots
```{r}
data(airquality)
plot(airquality$Temp, airquality$Ozone,
     xlab="Temperature", ylab="Ozone",
     main="Airquality relationship?")
pairs(airquality)
# Create function to make histogram with density superimposed
panel.hist <- function(x, ...)  {
  # Set user coordinates of plotting region
  usr <- par("usr"); on.exit(par(usr)) # I don't know how to use par() and its arguments, nor on.exit()
  par(usr = c(usr[1:2], 0, 1.5))
  par(new=TRUE)                    # Do not start new plot
  hist(x, prob=TRUE, axes=FALSE, xlab="", ylab="", 
       main="", col="lightgray")
  lines(density(x, na.rm=TRUE))    # Add density curve
}
# Create function to compute and print R^2
panel.r2 <- function(x, y, digits=2, cex.cor, ...) {
  # Set user coordinates of plotting region
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- cor(x, y, use="complete.obs")**2  # Compute R^2
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  if(missing(cex.cor)) cex.cor <- 1/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~ Ozone + Temp + Wind + Solar.R, data=airquality, 
      lower.panel=panel.smooth, upper.panel=panel.r2, 
      diag.panel=panel.hist) # how to use pairs() and its arguments

```


### Create a histogram plot

```{r}
data(LakeHuron)
hist(LakeHuron)
# Use the breaks option to specify the number of bins
# regardless of the size of the dataset
# Remove title and make bars gray
hist(LakeHuron, xlab="Level of Lake Huron (in feet)",
     freq=FALSE, breaks=c(574, 577, 578, 581, 585), 
     col="gray", main=NULL) 
# ues should include the entire range of observations).If the bins are of unequal widths, then hist by default produces a relative frequency histogram. Setting the freq=FALSE option forces the hist function to produce relative frequency histograms even when the bin widths are identical
lines(density(LakeHuron)) # what parameter to set for the height of y-axis
```


### Create a boxplot

```{r}
data(ToothGrowth)
str(ToothGrowth)
teeth <- subset(ToothGrowth, supp=="VC") # Use only some data
teeth
boxplot(len ~ factor(dose), data=teeth)
# Horiz. boxplot with whiskers from minimum to maximum value
attach(ToothGrowth)                      # Use all data
boxplot(len, horizontal=TRUE, range=0) # The standard boxplot where the whiskers extend to the minimum and maximum value can be obtained by setting the range option to zero.
```

### Create a bar plot

```{r}
m <- matrix(c(23, 20, 12, 5, 6, 9), ncol=3)  # Input data
relfrq <- prop.table(m, margin=1)            # The prop.table function computes the relative frequencies of a matrix given as the first argument. The second argument to prop.table determines if the proportions are relative to the row sums (margin=1) or the column sums (margin=2).
relfrq
t(relfrq)
# Make juxtaposed barplot
barplot(relfrq,  beside=TRUE, 
        legend.text=c("Hadsten", "Hammel"),
        names=c("<0.01", "0.01-0.10", ">0.10")) # beside  a logical value. If FALSE, the columns of height are portrayed as stacked bars, and if TRUE the columns are portrayed as juxtaposed bars
# Stacked relative barplot with labels added
barplot(relfrq,  beside=FALSE, 
        legend.text=c("Hadsten", "Hammel"),
        names=c("<0.01", "0.01-0.10", ">0.10"))
barplot(t(relfrq), names=c("Hadsten", "Hammel"), horiz=TRUE) # default is beside = FALSE
```

### Create a bar plot with error bars

```{r}
library(gplots)
data(airquality)
# Compute means and standard deviations for each month
mean.values <- by(airquality$Temp, airquality$Month, mean) # by() apply functions to specific vector
mean.values
sd.values <- by(airquality$Temp, airquality$Month, sd)
sd
barplot2(mean.values,
         xlab="Month", ylab="Temperature (F)",
         plot.ci=TRUE,
         ci.u = mean.values+1.96*sd.values,
         ci.l = mean.values-1.96*sd.values)
```  

> The barplot2 function from the gplots package can be used to create a bar plot with error bars. barplot2 is an enhancement of the standard barplot function that can plot error bars for each bar.
The function takes the same arguments as barplot but in addition you can specify the lower and upper limits for the individual error bars with the ci.l and ci.u options, respectively.


```{r}
airquality$Month <- factor(airquality$Month)
model <- lm(Temp ~ Month - 1, data=airquality) # look detail in statistical R
mean.values <- coef(summary(model))[,1]
mean.values
sem <- coef(summary(model))[,2]
scale <- qt(0.975, df=summary(model)$df[2])
barplot2(mean.values,
         xlab="Month", ylab="Temperature (F)",
         plot.ci=TRUE,
         ci.u = mean.values+scale*sem,
         ci.l = mean.values)
```


### Create a plot with estimates and confidence intervals

```{r}
library(gplots)
data(OrchardSprays)
fit <- lm(decrease ~ treatment - 1, data=OrchardSprays)
summary
```

> Sometimes it is desirable to present the results from an analysis graphically, for example by illustrating parameter estimates and their corresponding confidence intervals. These graphs can be created using standard plotting functions like plot and arrows, but the plotCI function from gplots package combines the necessary steps into a single function.
> plotCI requires one or two arguments, x and y which define the coordinates for the center of the error bars. y defaults to the vector 1:n if it is omitted. The uiw and liw options set the widths of the upper/right and lower/left error bar, respectively, and if liw is not specified it defaults to the same value as for uiw

```{r}
plotCI(coef(summary(fit))[,1], 
       uiw=qt(.975, df=56)*coef(summary(fit))[,2],
       ylab="Estimates and 95% confidence intervals", xaxt="n")
axis(1, 1:length(rownames(coef(summary(fit)))), 
     rownames(coef(summary(fit)))) # ?axis for arguments usage
```

> uiw  width of the upper or right error bar. Set to NULL or NA to omit upper bars.

> liw   width of the lower or left error bar. Defaults to same value as uiw. Set to NULL or NA to omit lower bars.


### Create a pyramid plot

```{r}

library(plotrix)
men <- c(334700, 357834, 328545, 367201, 411689, 
         359363, 336841, 178422, 72296, 9552, 139, 0)
women <- c(318662, 340357, 320462, 365180, 401521, 
           357405, 345884, 208057, 117843, 27914, 761, 0)
groups <- paste(seq(0, 110, 10), "-", seq(9, 119, 10), sep="")
groups                             # Show group labels
men <- 100*men/sum(men)            # Normalize men
women <- 100*women/sum(women)      # and women
men
women
pyramid.plot(men, women, labels=groups,
            lxcol="lightgray", rxcol="lightgray")
library(MASS)
data(survey)               
# Tabulate exercise and smoking status by gender
result <- by(survey, survey$Sex, 
             function(x) {table(x$Smoke, x$Exer)})
result
pyramid.plot(lx=result$Male, rx=result$Female,
             labels=row.names(result$Male),
             gap=10, unit="Number of students",
             top.labels=c("Male","Smoking Frequency","Female"),
             lxcol=c("darkgray", "white", "lightgray"),
             rxcol=c("darkgray", "white", "lightgray"))
```


Plot multiple series
Make a 2D surface plot
Make a 3D surface plot
Plot a 3D scatter plot
Create a heat map plot
Plot a correlation matrix
Make a quantile-quantile plot
Graphical model validation for linear models
More advanced graphics
Create a broken axis to indicate discontinuity
Create a plot with two y-axes
Rotate axis labels
Multiple plots
Add a legend to a plot
Add a table to a plot
Label points in a scatter plot
Identify points in a scatter plot
Visualize points, shapes, and surfaces in 3D and interact with them in real-time
Working with graphics
Exporting graphics
Produce graphics output in LaTeX-ready format
Embed fonts in postscript or pdf graphics