---
title: "R more dimensions dummies"
author: "Kenny"
date: "Sunday, January 25, 2015"
output: html_document
---
 
# Adding a Second Dimension
 
## Discovering a new dimension
 
### Creating your first matrix


```{r}

first.matrix <- matrix(1:12, ncol=4)
first.matrix
matrix(1:12, ncol=4, byrow=TRUE)

```

### Looking at the properties
 
```{r}

str(first.matrix)
dim(first.matrix)
length(first.matrix)
```



## Combining vectors into a matrix

```{r}

baskets.of.Granny <- c(12,4,5,6,9,3)
baskets.of.Geraldine <- c(5,4,2,4,12,9)
baskets.team <- rbind(baskets.of.Granny, baskets.of.Geraldine)
 
baskets.team
 
cbind(1:3, 4:6, matrix(7:12, ncol=2))
 
```

## How to use attributes

```{r}
baskets.team
# attributes(my.array)
attr(baskets.team,'season') <- '2010-2011'
attr(baskets.team,'season')
attr(baskets.team,'season') <- NULL
 
```

# Using the Indices
 
## Extracting values from a matrix
 
### Using numeric indices

```{r}

first.matrix[1:2, 2:3]
first.matrix[2:3,]

```
 
### Dropping values using negative indices

```{r}

first.matrix[-2,-3]
 
nr <- nrow(first.matrix)
id <- nr*2+2
first.matrix[-id]
 
first.matrix[-(2 * nrow(first.matrix) + 2)]
 
```

 
### Juggling dimensions  

> You can force R to keep all dimensions by using the extra argument drop from the indexing function. To get the second row returned as a matrix, you do the following:

```{r}

first.matrix[-c(1, 3), ]
first.matrix[2, , drop=FALSE]
 
```

## Replacing values in a matrix

```{r}


first.matrix[3, 2] <- 4
first.matrix
 
first.matrix[2, ] <- c(1,3)
first.matrix
 
first.matrix[1:2, 3:4] <- c(8,4,2,1)
first.matrix

```
 
# Naming Matrix Rows and Columns
 
## Changing the row and column names

```{r}

rownames(baskets.team) <- c('Granny','Geraldine')
rownames(baskets.team)
colnames(baskets.team) <- c('1st','2nd','3th','4th','5th','6th')
baskets.team
 
colnames(baskets.team)[3] <- '3rd'  # [] and () still confuses me sometimes
 
baskets.copy <- baskets.team
colnames(baskets.copy) <- NULL
baskets.copy
 
```

## Using names as indices

```{r}

baskets.team[, c("2nd","5th")]
 
baskets.team["Granny",]

```
 
# Calculating with Matrices
 
## Using standard operations with matrices

```{r}

first.matrix + 4
 
second.matrix <- matrix(1:3, nrow=3, ncol=4)
 
first.matrix + second.matrix
 

# first.matrix + second.matrix[,1:3] # gives error for illustration
# Error in first.matrix + second.matrix[, 1:3] : non-conformable arrays
 
first.matrix + 1:3
 
## Calculating row and column summaries
 
rowSums(baskets.team)
 
## Doing matrix arithmetic
 
### Transposing a matrix
first.matrix
t(first.matrix)
 
t(1:10)
first.matrix[2,]
t(first.matrix[2,]) # the way it is structured is interesting

```
 
# Inverting a matrix  

> no idea how it work

```{r}

square.matrix <- matrix(c(1,0,3,2,2,4,3,2,1),ncol=3)

solve(square.matrix)

```

 
### Multiplying two matrices

```{r}


first.matrix %*% t(second.matrix) # no idea how it works 
 
first.matrix %*% 1:4

1:3 %*% first.matrix
 
```

# Adding More Dimensions
 
## Creating an array
 
### Using the creator functions

```{r}

my.array <- array(1:24, dim=c(3,4,2))

my.array
 
```

### Changing the dimensions of a vector
 
```{r}

my.vector <- 1:24
dim(my.vector) <- c(3,4,2)
identical(my.array, my.vector)
 
```

## Using dimensions to extract values

```{r}


my.array[2,3,1]
 
my.array[, 3, 2, drop=FALSE]
 
 
my.array[2, , ]
  
```


# Combining Different Types of Values in a Data Frame
 
## Creating a data frame from a matrix
 
### Using the function as.data.frame

```{r}

baskets.df <- as.data.frame(t(baskets.team))

### Looking at the structure of a data frame


baskets.df
str(baskets.df)

### Counting values and variables
 
nrow(baskets.df)
length(baskets.df)

```

 
 
 
## Creating a data frame from scratch
 
### Making a data frame from vectors

```{r}

employee <- c('John Doe','Peter Gynn','Jolie Hope')
salary <- c(21000, 23400, 26800)
startdate <- as.Date(c('2010-11-1','2008-3-25','2007-3-14'))
 
employ.data <- data.frame(employee, salary, startdate)
 
str(employ.data)
 

```

### Keeping characters as characters

```{r}

employ.data <- data.frame(employee, salary, startdate, stringsAsFactors=FALSE)
str(employ.data)

```
 
## Naming variables and observations
 
### Working with variable names


```{r}

colnames(employ.data)
names(employ.data)
 
names(employ.data)[3] <- 'firstday'
names(employ.data)

```

### Naming observations

```{r}

rownames(employ.data)
rownames(employ.data) <- c('Chef','BigChef','BiggerChef')
employ.data
 
```

# Manipulating Values in a Data Frame
 
## Extracting variables, observations, and values
 
### Pretending it's a matrix

```{r}
baskets.df
baskets.df['3rd', 'Geraldine']
baskets.df[, 1]
baskets.df[, 1, drop=FALSE]

str(baskets.df[, 1, drop=FALSE])
 
```

### Putting your dollar where your data is
 
baskets.df$Granny
 
## Adding observations to a data frame
 
### Adding a single observation

```{r}

result <- rbind(baskets.df, c(7,4))
result
 
baskets.df <- rbind(baskets.df,'7th' = c(7,4))
baskets.df
 

```

### Adding a series of new observations using rbind
 
```{r}

new.baskets <- data.frame(Granny=c(3,8),Geraldine=c(9,4))
rownames(new.baskets) <- c('8th','9th')
baskets.df <- rbind(baskets.df, new.baskets)
baskets.df

```
 
### Adding a series of values using indices

```{r}

baskets.df[c('8th','9th'), ] <- matrix(c(3,8,9,4), ncol=2)
baskets.df
baskets.df[c('8th','9th'), ] <- c(3,8,9,4)
baskets.df
```

## Adding variables to a data frame
 
### Adding a single variable

```{r}

baskets.of.Gabrielle <- c(11,5,6,7,3,12,4,5,9)
baskets.df$Gabrielle <- baskets.of.Gabrielle
 
head(baskets.df, 4)
baskets.df
 

```

### Adding multiple variables using cbind

```{r}


new.df <- data.frame(
   Gertrude  =  c(3,5,2,1,NA,3,1,1,4),
   Guinevere =  c(6,9,7,3,3,6,2,10,6)
)
 
head(cbind(baskets.df, new.df), 4)
 
```

# Combining Different Objects in a List
 
## Creating a list
 
### Creating an unnamed list

```{r}

baskets.list <- list(baskets.team, '2010-2011')
baskets.list

```
 
### Creating a named list

```{r}

baskets.nlist <- list(scores=baskets.team, season='2010-2011')
baskets.nlist


### Playing with the names of elements

names(baskets.nlist)
 
### Getting the number of elements
 
length(baskets.list)
 
## Extracting elements from lists
 
### Using [[]]
 
baskets.list[[1]]
baskets.nlist[['scores']]
 
### Using []
 
baskets.list[-1]
baskets.nlist[names(baskets.nlist)=='season']

```
 
 
## Changing the elements in lists
 
### Changing the value of elements

```{r}

baskets.nlist[[1]] <- baskets.df
baskets.nlist[['scores']] <- baskets.df
baskets.nlist$scores <- baskets.df
 
baskets.nlist[1] <- list(baskets.df)
 
baskets.list[1:2] <- list(baskets.df, '2009-2010')

```
 
### Removing elements

```{r}


baskets.nlist[[1]] <- NULL
baskets.nlist$scores <- NULL
baskets.nlist['scores'] <- NULL
 
baskets.nlist <- list(scores=baskets.df, season='2010-2011')
baskets.nlist['scores'] <- list(NULL)
baskets.nlist
 
```
### Adding extra elements using indices

```{r}

baskets.nlist$players <- c('Granny','Geraldine')
baskets.nlist[['players']] <- c('Granny','Geraldine')
baskets.nlist['players'] <- list(c('Granny','Geraldine'))
 
baskets.list[[3]] <- c('Granny','Geraldine')
baskets.list[3] <- list(c('Granny','Geraldine'))

```
 
### Combining lists
 
```{r}
 
baskets.list <- list(baskets.team,'2010-2011')
players <- list(rownames(baskets.team))
 
c(baskets.list, players)

```

 
## Reading the output of str() for lists
 
```{r}

str(baskets.list)
```
