---
title: "R over excel dummies"
author: "Kenny"
date: "Saturday, January 24, 2015"
output: html_document
---

 
# Adding Row and Column Totals

```{r}

str(iris)
head(iris)
iris.num <- iris[, -5]
 
colSums(iris.num)
colMeans(iris.num)
 
apply(iris.num, 2, min)
apply(iris.num, 2, max)
 
sapply(iris.num, min)
sapply(iris.num, max)
 

```
 
# Formatting Numbers

```{r}

format(12345.6789, digits=9, decimal.mark=",",
   big.mark=" ",small.mark=".", , small.interval=3)
 
x <- colMeans(mtcars[, 1:4])
format(x, digits=2, nsmall=2)

```

>  %.1f means to format the first supplied value as a fixed point value with one digit after the decimal point, and %% is a literal that means print a %.

> the literal %3.2f means to format the value as a fixed point value with three digits before the decimal and two digits after the decimal

```{r}

x <- seq(0.5, 0.55, 0.01)
x
sprintf("%.1f %%", 100*x)
 
set.seed(1)
x <- 1000*runif(5)
x
sprintf("$ %3.2f", x)
 
```

> R cycles through the elements and places them into the sprintf() literals. Thus, %s (indicating format the value as a string) gets the value “bread” the first time and “cookies” the second time.
 
> You can do everything with paste() and format() that you can do with sprintf(), so you don’t really ever need to use it. But when you do, it can simplify your code.

```{r}

stuff <- c("bread", "cookies")
price <- c(2.1, 4)
sprintf("%s costed $ %3.2f ", stuff, price)
 
```


# Sorting Data

```{r}

head(mtcars)
with(mtcars, mtcars[order(hp), ])
with(mtcars, mtcars[order(hp, decreasing=TRUE), ])
 
```

# Making Choices with If

```{r}

mtcars <- within(mtcars,
   mpgClass <- ifelse(mpg < mean(mpg), "Low", "High"))
 
mtcars[mtcars$mpgClass == "High", ]
 

```


 
# Calculating Conditional Totals

```{r}

with(mtcars, mean(mpg))
with(mtcars, mean(mpg[hp < 150]))
with(mtcars, mean(mpg[hp >= 150]))
with(mtcars, length(mpg[hp > 150]))
 
```


 
# Transposing Columns or Rows

```{r}

x <- matrix(1:12, ncol=3)
x
t(x)
 
t(mtcars[1:4, ])
 
```

 
# Finding Unique or Duplicated Values

```{r}

unique(mtcars$cyl)
dupes <- duplicated(iris)
head(dupes)
which(dupes)
iris[dupes, ]
iris[!dupes, ]
nrow(iris[!dupes, ])
 
```

 
# Working with Lookup Tables

```{r}

index <- match("Toyota Corolla", rownames(mtcars))
index
mtcars[index, 1:4]
 

```
 
# Working with Pivot Tables

```{r}

with(mtcars, tapply(hp, list(cyl, gear), mean))

aggregate(hp~cyl+gear+am, mtcars, mean)
 
```

 
# Using the Goal Seek and Solver

```{r}

sales <- function(price) { 100 - 0.5 * price }
revenue <- function(price) { price * sales(price) }
 
 
par(mfrow=c(1, 2))
curve(sales, from=50, to=150, xname="price", ylab="Sales", main="Sales")
curve(revenue, from=50, to=150, xname="price", ylab="Revenue", main="Revenue")
par(mfrow=c(1, 1))
 
optimize(revenue, interval=c(50, 150), maximum=TRUE)
```
