---
title: "Basic concepts"
author: "Kenny"
date: "Friday, January 09, 2015"
output: html_document
---

### What is vector

> all vector shall have the elements with same type

> sequence vectors
```{r}
5:9
seq(5, 9)
seq(5, 9, 0.5)
```

### retrieve values from a vector

```{r}
sentence <- c('walk', 'the', 'plank')
sentence[3]

sentence[1]  # Many languages start array indices at 0, but R's vector indices start at 1
sentence[3] <- "dog" # assign new values within an existing vector

sentence[4] <- 'to'

sentence[c(1, 3)] # use a vector within the square brackets to access multiple values

sentence[2:4]  # Get the second through fourth words:

sentence[5:7] <- c('the', 'poop', 'deck')

sentence[6]
```

### assign names to vector's elements

```{r}
ranks <- 1:3
names(ranks) <- c("first", "second", "third") # assign names to elements of a vector
```

### retrieve elements from vector by names

```{r}
ranks # still vector only with names
ranks["first"] # still a value element of a vector only with a name
ranks["third"] <- 4 # assign a different value to the element with name "third"
```

### barplot to draw vector

```{r}
barplot(1:100) # draw bars from vector
vesselsSunk <- c(4, 5, 1)
barplot(vesselsSunk)
```

> names of vector elements will show on graphs

```{r}
names(vesselsSunk) <- c("England", "France", "Norway")
barplot(vesselsSunk)

```

### Vector math

```{r}
a <- c(1, 2, 3)
a + 1
a == c(1, 99, 3) # each element compare to each element of different vector
sin(a) # function apply to each element

```

### How to plot two vectors 
```{r}
x <- seq(1, 20, 0.1)
y <- sin(x)
plot(x, y)
```

### How to deal with NA
```{r}
a <- c(1, 3, NA, 7, 9)
sum(a)
sum(a, na.rm = TRUE)
```

### Difference between matrix and array or vector

> A matrix is just a fancy term for a 2-dimensional array

### Build a matrix

```{r}
matrix(0, 3, 4) # 3 high, 4 wide, with 0 as value

a <- 1:12
print(a) # prove a is only a vector
matrix(a, 3, 4) # a vector is turned into a matrix

plank <- 1:8 # a vector or one dimensional array
dim(plank) <- c(2, 4) # give this vector(one dim of array) a two dimension attribute
print(plank) # automatically, the vector is turned into 2-dimension array = matrix

matrix(8, 5, 5) # make matrix with matrix() function, it can be 8 or a vector
plank <- matrix(1:8, 2, 4, byrow = FALSE)
print(plank)
```

### retrieve elements from Matrix
```{r}
plank[2, 3]  # as matrix is 2 dim, so two number to locate value
plank[1, 4]
plank[1, 4] <- 0
plank[2,]
plank[, 4]
plank[, 2:4]
```

### matrix plotting

> Text output is only useful when matrices are small. When working with more complex data, we need matrix plotting  

> Let's draw an elevation map of a sandy beach, 1 meter above sea level

```{r}
elevation <- matrix(1, 10, 10) # we think of a 3D sandy beach 
elevation[4, 6] <- 0 # we dig down 1 meter at a specific place on the beach
contour(elevation) # draw a contour map
persp(elevation) # draw a 3D map, but persp automatically expands the view so that your highest value (the beach surface) is at the very top
persp(elevation, expand=0.2) # more reasonable graph

```
> a 3D map of a dormant New Zealand volcano

```{r}
contour(volcano)
persp(volcano, expand=0.2)
image(volcano)
```


### summary statistics on mean

```{r}
limbs <- c(4, 3, 4, 3, 2, 4, 4, 4) # collect data about crew members
names(limbs) <- c('One-Eye', 'Peg-Leg', 'Smitty', 'Hook', 'Scooter', 'Dan', 'Mikey', 'Blackbeard')
mean(limbs)
barplot(limbs)
abline(h = mean(limbs)) # draw a horizontal line display mean
```

### graph ablines differ in sd

```{r}
pounds <- c(45000, 50000, 35000, 40000, 35000, 45000, 10000, 15000)
barplot(pounds)
meanValue <- mean(pounds)
abline(h = meanValue) # line on mean value
deviation <- sd(pounds) # value of 1 sd
abline(h = meanValue + deviation) # line on mean + 1sd
abline(h = meanValue - deviation)
```


### Create factors
> Often your data needs to be grouped by category = factors

```{r}
chests <- c('gold', 'silver', 'gems', 'gold', 'gems')
types <- factor(chests) # character vector turned into factor, no more quotation marks
print(chests) # chests is still the same vector
print(types) # types is the vector chests with factor or levels attribute

as.integer(types) # values of types are not strings but integer references to one of the factor's levels
levels(types) # but levels themselves are strings or characters, so with quotations
```

### graph factors with pch  
> graph the chests by weight and value  

```{r}
weights <- c(300, 200, 100, 250, 150)
prices <- c(9000, 5000, 12000, 7500, 18000)
plot(weights, prices) # both weights and prices correspond to chests or types

plot(weights, prices, pch=as.integer(types)) # converting the factor to integers, and passing it to the pch argument

legend("topright", c("gems", "gold", "silver"), pch=1:3) # make legends to correspond to pch

legend("topright", levels(types), pch=1:length(levels(types))) # make it automatic given future changes
```



### Why full control of graphics in R can be frustrating

> **There are many graphics systems in R, and they don’t necessarily play well with each other**





**Base graphics**  

* is the oldest and most basic. Base graphics produce static pictures        
* its par function is used to set up the graphics parameters that control the layout and look of the graphics  

**grid**  

* is a much more flexible system of graphics.   
* there are graph objects (see More R key objects for the idea of “object”).  

**lattice**  

* contains a set of graphics techniques that are very useful for data analysis.

**ggplot2**  

* is a plotting system for R, trying to take the good parts of base and lattice graphics  
* It takes care of many fiddly details that make plotting a hassle (like drawing legends)  
* provides a powerful model of graphics that makes it easy to produce complex multi-layered graphics.  

**Useful links**

* Cran Task View on [graphics](http://cran.r-project.org/web/views/Graphics.html)

### how many different objects in R?

**What is an object?**

* assign a name, if you can, it is; if you cannot assign a name to it, it is not
* something that we can hold, turn upside down, and look at

**What is Attribute?**

* Attributes are spice.
* objects have a main part, but can have one or more attributes
* same bowl of rice and beans, with different spice, they are different dishes.
* name is attribute: 
    * The elements of an atomic vector can each have a name. 
    * The components of a list can each have a name.
* class is a very important attribute.

**What are inherent characteristics of objects?**

* Objects have a “length”. 
    * The length of an atomic vector is the number of elements it has. 
    * The length of a list is the number of components that it has. 
    * The length of NULL is zero.

* Objects have a “mode”. 
    * This says what kind of object they are. 
    * mode() function tell you the mode of an object. 
    * typeof() function tell slightly more specific about an object.

**What are Atomic Types?**

* There are three atomic types  
    * Numeric objects hold numbers.
    * Logical objects have values that are TRUE, FALSE and NA.
    * Character objects have a string as each element.

**What are Matrices and Data Frames?**

* similarities 
    * matrices and data frames are rectangular objects with rows and columns.
    * Both are linear structures pretending to be rectangular. 
    * Both have names for the rows and the columns.
* differences
    * A matrix is a vector that has a dim attribute. 
    * The dim is a vector of two integers saying how many rows and columns there are. 
    * A data frame has a class attribute that is "data.frame". 
    * It is really a list with as many components as there are columns. 
    * Each component has to have the same number of elements (the number of rows).
    * check with function is.data.frame(x)

**What are Factors?**

* Factors have two key attributes: 
  
    * a class attribute and a levels attribute. 
    * The levels is a character vector that gives the possible categories for the object. 
    * ???The basic part of the object is a vector of integers that are the location of the category in the levels vector.


**Missing values**

* All of the atomic modes have a missing value. 
    * This is printed as NA.
    * test for missing values with the is.na function. For example: is.na(x)



**See inside structure of object**

```{r}
examp <- list(A=1:10, B=letters, C=list(NULL, TRUE))
str(examp)

```

**Function are objects too**


### What 2.3e6 and 2.3e-4 mean?

* 2.3e6 = 2.3 with decimal point moving right 6
* 2.3e6 = 2.3 with decimal point moving left 6

### Watch for integer and floating point

* they differ when floating point is result of rounding
```{r}
seq(0, 1, by = .1)
seq(0, 1, by = .1)[4] # the result is rounding outcome, not the real outcome
seq(0, 1, by = .1)[4] == .3 # therefore it is false not true here
seq(0, 1, by = .1)[4] - .3 
print(seq(0, 1, by=.1)[4], digits=20)

```

### How to replace . with ,

**replace . with , by format(, decimal.mark=)**
```{r}
format(seq(0, 1, by=.5), decimal.mark=",") # change to , by format(x, decimal.mark = ",")

```

**replace . with , by option at beginning**
```{r}
options(OutDec=",")
seq(0, 1, by=.5)
```

**use sub() to replace . with ,**
```{r}
nums <- format(seq(0, 1, by=.5), decimal.mark=",") # , makes number into strings
nums
sub(",", ".", nums) # sub() work with strings?
as.numeric(sub(",", ".", nums)) # turn string into numbers again
```

**replace multiple , with nothing**
```{r}
as.numeric(gsub(",", "", "2,300,000"))
```

### How R understand infinity and not-a-number

**What R's limit on infinit**

```{r}
1e308
1e308 * 10
```

**What R consider to be NaN**

```{r}
0/0
Inf - Inf
```

### Understand integer and floating point with examples

**Difference between mode and typeof, numberic and double**

```{r}
mode(1)
typeof(1)
mode(1:2)
typeof(1:2)

is.integer(1)
is.integer(1:2)
is.integer(as.integer(1))
is.integer(as.double(1:2))

is.integer(1L)
```

