---
title: "R Controlling logical flow"
author: "Kenny"
date: "Monday, January 12, 2015"
output: html_document
---
### Making Choices with if Statements

> Instead of using a classic loop structure, you use a function to apply another function on any of the objects. This way of looping over values is one of the features that distinguishes R from many other programming languages.

```{r}
 
priceCalculator <- function(hours, pph=40){  # this is proper function
    net.price <- hours * pphs 
    round(net.price)
}
 
priceCalculator <- function(hours, pph=40){
    net.price <- hours * pph
    if(hours > 100) {                        # same function include if to make a choice
      net.price <- net.price * 0.9
    }
    round(net.price)
}
 
priceCalculator(hours = 55)                # run the function with normal data
priceCalculator(hours = 110)               # run function with data specified in the choice
 
priceCalculator <- function(hours, pph=40){
    net.price <- hours * pph
    if(hours > 100) net.price <- net.price * 0.9   # same function with specific choice but if statment in simplified version without {}
    round(net.price)
}
 

```

### When to use if ... else

> In some cases, you need your function to do something if a condition is true and something else if it is not.

> if(public == TRUE) is about as redundant as asking if white snow is white, if(public) is good coding

```{r}

priceCalculator <- function(hours, pph=40, public=TRUE){ # hours = necessity data, pph = constant, 
                                                         # public = TRUE default setting and arg for if
                                                         # ...else situation
    net.price <- hours * pph
    if(hours > 100) net.price <- net.price * 0.9         # a free choice
    if(public) {                                         # yes choice from yes or no situation
      tot.price <- net.price * 1.06
    } else {                                             # no choice from yes or no situation
      tot.price <- net.price * 1.12
    }
    round(tot.price)
}
 
priceCalculator(25,public=TRUE)                          # 25 stands for hours, public = TRUE default 
priceCalculator(25,public=FALSE)                         # value no need mentioning, unless it is not 
                                                         # default but public = FALSE
```


> **make if...else statement shorter**
>  Putting the else statement at the end of a line and not the beginning of the next one is a good idea. 
>  If you put else at the beginning of the second line, R considers the first line finished and complains. 

```{r}

priceCalculator <- function(hours, pph=40, public=TRUE){
    net.price <- hours * pph
    if(hours > 100) net.price <- net.price * 0.9         # A short way of writing if statement
    if(public) tot.price <- net.price * 1.06 else        # a short way of writing if...else statement
               tot.price <- net.price * 1.12
    round(tot.price)
}

```
 
 
> **to make if...else statement even shorter**
```{r}

priceCalculator <- function(hours, pph=40, public=TRUE){
    net.price <- hours * pph
    if(hours > 100) net.price <- net.price * 0.9
    tot.price <- net.price * if(public) 1.06 else 1.12   # R will access if...else statement first
    round(tot.price)
}

```
 
### taking a vector of values together automatically not one value a time --Vectorizing Choices
 
> current function can only take one value at a time, but even two values together

```{r}
priceCalculator
priceCalculator(c(25,110))  # the problem is current function can only take one value not two
priceCalculator(110)
c(25, 110) > 100            # both values have different criteria, one < 100, the other > 100
                  
```

> if...else: yes and no for doing different things; ifelse: yes and no for selecting diff values  
> test for true or false on more than one value, there need more choices for selections

```{r}


ifelse(c(1,3) < 2.5, 1:2, 3:4) # the first value of second arg, the second value of third arg

ifelse(c(3,2) < 2.5, 1:2, 3:4) # the second value of second arg, the first value of third arg

```

> example with only one ifelse

```{r}

my.hours <- c(25,110)

my.hours * 40 * ifelse(my.hours > 100, 0.9, 1) # yes and no for selecting diff values

```
 
> function with two ifelse

```{r}

priceCalculator <- function(hours,pph=40,public){  # two args used in ifelse, pph = constant
    net.price <- hours * pph
    net.price <- net.price * ifelse(hours > 100 , 0.9, 1) # ifelse on hours
    tot.price <- net.price * ifelse(public, 1.06, 1.12)   # ifelse on public
    round(tot.price)
}
 
clients <- data.frame(                 # two args formulate features of each client
  hours = c(25, 110, 125, 40),         # There are four clients = four diff set of (two) args
  public = c(TRUE,TRUE,FALSE,FALSE)
)
 
with(clients, priceCalculator(hours, public = public))  # apply the data to function 
 

```

### What if there are Multiple Choices to choose from
 
> **Chaining/nesting if...else situation**

> three options: private, public and foreign, with three diff rates
 

> **most formal and lengthy way of coding**

```
if(client=='private'){
 tot.price <- net.price * 1.12      # 12% VAT
} else {
 if(client=='public'){
   tot.price <- net.price * 1.06    # 6% VAT
 } else {
   tot.price <- net.price * 1    # 0% VAT
 }
}
```

> **saved one pair of curly braces from else**

```
if(client=='private'){
   tot.price <- net.price * 1.12
} else if(client=='public'){          # one pair is saved here
   tot.price <- net.price * 1.06
} else {
   tot.price <- net.price
}
```

> **most efficient way of coding with nesting ifelse twice** 

```
VAT <- ifelse(client=='private', 1.12,           # lots of curly braces saved
         ifelse(client == 'public', 1.06, 1)
      )
tot.price <- net.price * VAT

```
> **chained if...else statements can only on one value at a time, chained ifelse can work on vectors  
> **if there are more than two ifelse, then switch is needed


### Switching between possibilities
 
### Making choices with switch
 
# Code example # NOT run
# VAT <- switch(client, private=1.12, public=1.06, abroad=1)
 
 
 
### Using default values in switch
 
# Code example # NOT run
# VAT <- switch(client, private=1.12, public=1.06, 1)
 
client <- 'other'
switch(client, private=1.12, public=1.06, 1)
 
 
# Looping Through Values
 
## Constructing a for loop
 
## Calculating values in a for loop
 
### Using the values of the vector
 
priceCalculator <- function(hours, pph=40, client){
    net.price <- hours * pph *
                   ifelse(hours > 100, 0.9, 1)
 
    VAT <- numeric(0)
    for(i in client){
      VAT <- c(VAT,switch(i, private=1.12, public=1.06, 1))
    }
 
    tot.price <- net.price * VAT
    round(tot.price)
}
 
 
clients$type <- c('public','abroad','private','abroad')
priceCalculator(clients$hours, client=clients$type)
 
### Using loops and indices
 
nclient <- length(client)
VAT <- numeric(nclient)
for(i in seq_along(client)){
  VAT[i] <- switch(client[i], private=1.12, public=1.06, 1)
}
VAT
 
# Looping without Loops: Meeting the Apply Family
 
songline <- 'Get out of my dreams...'
for(songline in 1:5) print('...Get into my car!')
 
songline
 
## Looking at the family features
 
## Meeting three of the members
 
 
## Applying functions on rows and columns
 
### Counting birds
 
counts <- matrix(c(3,2,4,6,5,1,8,6,1), ncol=3)
colnames(counts) <- c('sparrow','dove','crow')
counts
 
apply(counts, 2, max)
 
### Adding extra arguments
 
counts[2, 2] <- NA
apply(counts,2,max)
apply(counts, 2, max, na.rm=TRUE)
 
## Applying functions to listlike objects
 
### Applying a function to a vector
 
#### Using switch on vectors
 
sapply(c('a','b'), switch, a='Hello', b='Goodbye')
 
#### Replacing a complete for loop with a single statement
 
priceCalculator <- function(hours, pph=40, client){
  net.price <- hours * pph * ifelse(hours > 100, 0.9, 1)
 
  VAT <- sapply(client, switch, private=1.12, public=1.06, 1)
 
  tot.price <- net.price * VAT
  round(tot.price)
}
 
### Applying a function to a data frame
 
sapply(clients,class)
 
### Simplifying results (or not) with sapply
 
sapply(clients, unique)
 
### Getting lists using lapply
 
sapply(clients[c(1,3), ], unique)
 
lapply(clients[c(1,3), ], unique)